---
title: "Team Valuation Project"
author: "Terrick Boire"
date: "2023-05-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(lubridate)
library(caret)
library(stats)
library(ggplot2)
library(corrplot)
library(dplyr)
```


# Loading the Data into R

```{r}
franchise.data <- read.csv("franchise values.csv")
franchise.data
```



### DATA CLEANING

# Checking the dimensions

```{r}
dim(franchise.data)
```
# Checking the structure of the data frame

```{r}
str(franchise.data)
```

# Dealing with Charaters that should be ints by year to not lose data


```{r}
franchise.2023 <- franchise.data %>%
  filter(Year == "2023")

franchise.2023
```

# Removing ",s" and making numeric

```{r}
franchise.2023$Revenue <- as.numeric(gsub(",","",franchise.2023$Revenue))

franchise.2023$Value <- as.numeric(gsub(",","",franchise.2023$Value))

franchise.2023$Year.over.Year.Value.Increase <- as.numeric(gsub(",","",franchise.2023$Year.over.Year.Value.Increase))

franchise.2023$Debt.Value <- as.numeric(gsub(",","",franchise.2023$Debt.Value))

franchise.2023$Income <- as.numeric(gsub(",","",franchise.2023$Income))

franchise.2023$Population <- as.numeric(gsub(",","",franchise.2023$Population))

franchise.2023
```



# Franchise values for 2021

```{r}
franchise.2021  <- franchise.data %>%
  filter(Year == "2021")
  
franchise.2021
```

# Removing the % symbol from data

```{r}
franchise.2021$Year.over.Year.Value.Increase <- gsub( "%", "", as.character(franchise.2021$Year.over.Year.Value.Increase))

franchise.2021$Debt.Value <- gsub( "%", "", as.character(franchise.2021$Debt.Value))

franchise.2021
```

# Removing ",s" from data and making numeric

```{r}
franchise.2021$Revenue <- as.numeric(gsub(",","",franchise.2021$Revenue))

franchise.2021$Value <- as.numeric(gsub(",","",franchise.2021$Value))

franchise.2021$Year.over.Year.Value.Increase <- as.numeric(gsub(",","",franchise.2021$Year.over.Year.Value.Increase))

franchise.2021$Debt.Value <- as.numeric(gsub(",","",franchise.2021$Debt.Value))

franchise.2021$Income <- as.numeric(gsub(",","",franchise.2021$Income))

franchise.2021$Population <- as.numeric(gsub(",","",franchise.2021$Population))

franchise.2021
```


# Converting data to reflect original % value
```{r}
franchise.2021$Year.over.Year.Value.Increase <- franchise.2021$Year.over.Year.Value.Increase / 100

franchise.2021
```


# Converting data to reflect original 5 value
```{r}
franchise.2021$Debt.Value <- franchise.2021$Debt.Value / 100

franchise.2021
```

# Franchise values for 2020


```{r}
franchise.2020  <- franchise.data %>%
  filter(Year == "2020")
  
franchise.2020
```

# Removing ",s" from 2020 data and making numeric

```{r}
franchise.2020$Revenue <- as.numeric(gsub(",","",franchise.2020$Revenue))

franchise.2020$Value <- as.numeric(gsub(",","",franchise.2020$Value))

franchise.2020$Year.over.Year.Value.Increase <- as.numeric(gsub(",","",franchise.2020$Year.over.Year.Value.Increase))

franchise.2020$Debt.Value <- as.numeric(gsub(",","",franchise.2020$Debt.Value))

franchise.2020$Income <- as.numeric(gsub(",","",franchise.2020$Income))

franchise.2020$Population <- as.numeric(gsub(",","",franchise.2020$Population))

franchise.2020
```


# Removing MLS data so that datasets are of equal length

```{r}
franchise.2020 <- franchise.2020 %>%
  filter(League != "MLS")

franchise.2020
```



# Adding population data for 2021 based on 2020 data

```{r}
franchise.2021$Population <- ifelse(franchise.2021$Team == franchise.2020$Team, franchise.2020$Population, 0)

franchise.2021
```


# Adding populations based on 2021 information for 2023

```{r}
franchise.2023 <- franchise.2023 %>% 
  mutate(Population = ifelse(Team %in% franchise.2021$Team, 
                           franchise.2021$Population[Team == franchise.2021$Team],0))

franchise.2023
```


# Adding population for 2023 data for new teams or teams with name changes

# Adding population for cle guardians formerly cle indians
```{r}
franchise.2023[8, "Population"] <- 2055612

franchise.2023
```

# Adding population for Wash Commanders previously wash football team

```{r}
franchise.2023[91, "Population"] <- 6131977

franchise.2023
```

# Adding population for new team seattle kraken (got population from google)

```{r}
franchise.2023[116, "Population"] <- 733919

franchise.2023
```

# Putting 0 in for year over year value, income, and debt value since it is the first year (For all 2020 data)


```{r}
franchise.2020 <- franchise.2020 %>% 
  mutate(Year.over.Year.Value.Increase = ifelse(is.na(Year.over.Year.Value.Increase), 0, Year.over.Year.Value.Increase),
                    Debt.Value = ifelse(is.na(Debt.Value), 0, Debt.Value),
                    Income = ifelse(is.na(Income), 0, Income))

franchise.2020
```


# Adding income for the 1 na remaining for 2021 calgary flames. Got income from https://www.statista.com/statistics/374824/calgary-flames-operating-income/

```{r}
franchise.2021[97, "Income"] <- -39000000

franchise.2021
```

# Replacing the unknow Yankees 2023 debt value with the 2021 debt value

```{r}
franchise.2023[19, "Debt.Value"] <- 0

franchise.2023
```



# Putting the cleaned data back together

```{r}

franchise.data <- bind_rows(franchise.2023, franchise.2021, franchise.2020)
```



```{r}
sum(is.na(franchise.data))
```


# Getting a summary of the data

```{r}
summary(franchise.data)
```

# Finding the correlation between all numeric variables

```{r}
cor_franchise <- franchise.data %>%
  select_if(is.numeric)

c <- round(cor(cor_franchise), digits = 2)
c
```
# Viewing the correlation plot of the variables

```{r}
corrplot(c)

```
# Seeing the variables with high correlation

```{r}
highCorr <- findCorrelation(c, cutoff = 0.5, names = TRUE)
highCorr
```



# Comparing Value and revenue from the  MLB to NBA

# Filtering the data to only use MLB and NBA data
```{r}
MLB <- franchise.data %>% 
  filter(League == "MLB") 

NBA <- franchise.data%>%
  filter(League == "NBA" )
```


# Combining MLB and NBA data
```{r}
mlb_nba <- bind_rows(MLB,NBA)
```

# Creating the scatterplot
```{r}
ggplot(mlb_nba, aes(x = Revenue, y = Value, color = League))+
  geom_point()+
  facet_wrap(~League)+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_continuous(labels = scales::dollar_format())+
  scale_y_continuous(labels = scales::dollar_format())
```



# Creating a boxplot to compare mlb and nba revenue and value

```{r}
ggplot(mlb_nba, aes(x = Revenue, y = Value, color = League))+
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_continuous(labels = scales::dollar_format())+
  scale_y_continuous(labels = scales::dollar_format())
```




# Value revenue MLB to NHL
# Filtering the data to only use MLB and NBA data
```{r}
NHL <- franchise.data%>%
  filter(League == "NHL" )


```

# Combining MLB and NHL Data
```{r}
mlb_nhl <- bind_rows(MLB, NHL)
```


# Creating a Scatterplot
```{r}
ggplot(mlb_nhl, aes(x = Revenue, y = Value, color = League))+
  geom_point()+
  facet_wrap(~League)+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_continuous(labels = scales::dollar_format())+
  scale_y_continuous(labels = scales::dollar_format())
```

# Creating a boxplot to compare mlb and nhl revenue and value

```{r}
ggplot(mlb_nhl, aes(x = Revenue, y = Value, color = League))+
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_continuous(labels = scales::dollar_format())+
  scale_y_continuous(labels = scales::dollar_format())
```




# Comparing the 2021 Seasons to other Seasons

# Filtering the data by year
```{r}
Season_21 <- franchise.data %>% 
  filter(Year == 2021) 

Season_23 <- franchise.data%>%
  filter(Year ==  2023)

Season_20 <- franchise.data%>%
  filter(Year == 2020)
```

```{r}
ggplot(franchise.data, aes(x = Revenue, y = Income, color = League))+
  geom_boxplot()+
  facet_wrap(~Year)+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_continuous(labels = scales::dollar_format())+
  scale_y_continuous(labels = scales::dollar_format())
```


# Creating a regression model to predic the future value

```{r}
fit <- lm(Value ~ Revenue + Income + Year.over.Year.Value.Increase, data = franchise.data)
fit_sum <- summary(fit)

```


```{r}
plot(fit)
```


# Making new data set To look at predicted values easier

```{r}
Predicted_regression <- bind_cols(franchise.data$Team, franchise.data$League, franchise.data$Year, franchise.data$Value, fit$fitted.values)

Predicted_regression
```


# Creating a Second Model (Regression Tree)



```{r}
library(tree)
```


```{r}
na.omit(franchise.data)

set.seed (1)
training <- sample (1: nrow (franchise.data), nrow (franchise.data) / 2)
tree.Values <- tree (Value ~ ., franchise.data , subset = training)
summary (tree.Values)

```


```{r}
plot (tree.Values)
text (tree.Values , pretty = 0)
```


# Pruning the tree to get better results

```{r}
prune.Values <- prune.tree (tree.Values , best = 5)
plot (prune.Values)
text (prune.Values , pretty = 0)
```



```{r}
yhat <- predict (tree.Values , newdata = franchise.data[-training , ])
Values.test <- franchise.data[-training , "Value"]
plot (yhat , Values.test)
abline (0, 1)
mean ((yhat - Values.test)^2)
```


# Predictind values from the tree
```{r}
yhat_predicted <- predict(tree.Values, newdata = franchise.data)
```



```{r}
Predicted_tree <- bind_cols(franchise.data$Team, franchise.data$League, franchise.data$Year, franchise.data$Value, yhat_predicted)
Predicted_tree
```


